Livin' on the edge: Precision yield data shows evidence of ecosystem services from field boundaries

### 2022-03-04 Manuscript/manuscript.Rmd
Updated manuscript

Precision yield data were collected directly from farmers across Alberta.
Farmers were solicited for yield data through local agronomists, and we received `r nrow(datSource)` field-years of data from `r length(unique(datSource$grower))` growers across a total of `r length(unique(datSource$year))` years (`r min(unique(datSource$year))`-`r max(unique(datSource$year))`).
We converted raw data to a standard csv format using Ag Leader SMS®.
`r round(100*sum(grepl('(Wheat|Canola|Peas)',datSource$crop))/nrow(datSource))`% of the crop types where either wheat (_Triticum aestivum_), canola (_Brassica napus_), or peas (_Pisum sativum_), three of the most common crops in rotation in Alberta [@statscan_canola2014].
The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`), canola (`r sum(datSourceFinal$crop=='Canola')`), or peas (`r sum(datSourceFinal$crop=='Peas')`), for a total of `r nrow(datSourceFinal)` field-years of data.
Individual fields ranged from `r round(min(fieldAreas),1)` to `r round(max(fieldAreas),1)` ha in area (mean: `r round(mean(fieldAreas),1)` ha), and contained between `r min(yearsPerField)` and `r max(yearsPerField)` years of data (mean: `r round(mean(yearsPerField),1)`), for a total of `r round(sum(datSourceFinal$npoints)/1e6,1)` million data points.

```{r fieldLocataions, echo=FALSE, fig.cap='Locations of fields that data were collected from', out.width='100%'}  
  include_graphics('../Figures/Field Locations.png',dpi=NA)
```

Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m$^2$), dry yield (tonnes), and the spatial location.
Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
Seeding and application rates were constant across fields, so we did not consider inputs in our analysis.
Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows [@griffin2007; @whelan2013], or when the combine is accelerating or decelerating [@arslan2002].
Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix A for R code):

@@

```{r boundaryTypes, echo=FALSE, fig.cap='Boxplot of field boundary compositions across sampled fields. The x-axis indicates the boundary type, while the y-axis indicates the percentage of all field boundaries covered by that boundary type.', out.width='100%'}

bLengths %>% 
  mutate(croptype=datSourceFinal$crop) %>% 
  mutate(location=ifelse(grepl('(Alvin|Dean)',filename),'Southern Alberta','Central Alberta')) %>%
  group_by(location) %>% mutate(nFields=length(unique(filename))) %>% ungroup() %>% 
  mutate(location=paste0(location,' (N = ',nFields,')')) %>%
  select(filename,croptype,location,contains('prop_')) %>% 
  pivot_longer(contains('prop_')) %>% 
  mutate(name=gsub('prop_','',name),value=value*100) %>% na.omit() %>%
  mutate(name=factor(name,labels=c('Bare','Grassland','Other\nCrop','Shelterbelt','Standard','Wetland'))) %>% 
  group_by(croptype) %>% mutate(croptype=paste0(croptype,'\n(N = ',length(unique(filename)),')')) %>% 
  ggplot(aes(x=name,y=value,fill=croptype))+geom_boxplot()+
  facet_wrap(~location,ncol=1) +
  labs(y='Percentage of field boundary',fill='Crop')+
  scale_fill_manual(values=c('grey50','grey80','white'))+
  theme(axis.title.x = element_blank(),legend.position = 'bottom')

# bLengthSummary %>% #Table version
#   kable(.,format='latex',caption='Mean and SD of proportion of boundary types across all fields, along with the number of fields bordered by at least some of the boundary type.',col.names = c('Type','Mean','SD','Number of Fields'),digits = c(1,4,4,1),escape = FALSE)

@@

>>>>

Precision yield data were collected directly from farmers across Alberta (Figure \@ref(fig:fieldLocations)).
Farmers were solicited for yield data through local agronomists, and we received `r nrow(datSource)` field-years of data from `r length(unique(datSource$grower))` growers across a total of `r length(unique(datSource$year))` years (`r min(unique(datSource$year))`-`r max(unique(datSource$year))`).
We converted raw data to a standard csv format using Ag Leader SMS®.
`r round(100*sum(grepl('(Wheat|Canola|Peas)',datSource$crop))/nrow(datSource))`% of the crop types where either wheat (_Triticum aestivum_), canola (_Brassica napus_), or peas (_Pisum sativum_), three of the most common crops in rotation in Alberta [@statscan_canola2014].
The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`), canola (`r sum(datSourceFinal$crop=='Canola')`), or peas (`r sum(datSourceFinal$crop=='Peas')`), for a total of `r nrow(datSourceFinal)` field-years of data.
Individual fields ranged from `r round(min(fieldAreas),1)` to `r round(max(fieldAreas),1)` ha in area (mean: `r round(mean(fieldAreas),1)` ha), and contained between `r min(yearsPerField)` and `r max(yearsPerField)` years of data (mean: `r round(mean(yearsPerField),1)`), for a total of `r round(sum(datSourceFinal$npoints)/1e6,1)` million data points.

Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m$^2$), dry yield (tonnes), and the spatial location.
Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
Seeding and application rates were constant across fields, so we did not consider inputs in our analysis.
Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows [@griffin2007; @whelan2013], or when the combine is accelerating or decelerating [@arslan2002].
Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix B for R code):

@@

@@

\newpage

# Figures {-}

\newpage

# Tables {-}
